---
title: "STAT 512 Final Group Project Rmd"
output: html_document
author: "Team 6 - Section 3"
---

```{r}
#Import Data
bball = read.csv("nba.games.stats.csv", header = TRUE, sep =",")

# colnames(bball)
# IND = bball[bball$Team == "IND",]
# dim(IND)
# Filter to just take one team "IND"
bball = bball[bball$Team == "IND",]

# Subsetting data
myvars = c("TeamPoints","X3PointShots.","FieldGoals.","Assists","TotalRebounds","Turnovers","TotalFouls","Steals")
bball = bball[myvars]
# colnames(bball)

# Summary
summary(bball)
dim(bball)

```


# Preliminary Analysis
```{r}
# Assign Valriables
Points = bball$TeamPoints
Assists = bball$Assists
ThreeP = bball$X3PointShots.
Rebounds =   bball$TotalRebounds
Turnovers= bball$Turnovers
FieldGoals = bball$FieldGoals
Fouls = bball$TotalFouls
Steals = bball$Steals

Project_data = data.frame(Points, Assists, ThreeP, Rebounds, Turnovers, FieldGoals, Fouls, Steals)
summary(Project_data)

par(mfrow=c(2,2))
#Assists and Team points
plot(Assists, Points, main = "Assists and Team Points", xlab ="Assists", ylab="Team Points")
mod1 <-lm(Points ~ Assists)
abline(mod1, col=3)
hist(Assists, main = "Histogram of Assists", xlab = "Assists")

#Rebounds and Team points
plot(Rebounds, Points, main = "Rebounds and Team Points", xlab ="Rebounds", ylab="Team Points")
mod2 <-lm(Points ~ Rebounds)
abline(mod2, col=3)
hist(Rebounds, main = "Histogram of Rebounds", xlab = "Rebounds")

#3 Point Shot percentage and Team points
plot(ThreeP, Points, main = "3 Point Shot Percentage and Team Points", xlab ="3 Point Shot Percentage", ylab="Team Points")
mod3 <-lm(Points ~ ThreeP)
abline(mod3, col=3)
hist(ThreeP, main = "Histogram of 3 Point shot percentage", xlab = "3 Point shot percentage")

#Turnovers and Team points
plot(Turnovers, Points, main = "Turnovers and Team Points", xlab ="Turnovers", ylab="Team Points")
mod4 <-lm(Points ~ Turnovers)
abline(mod4, col=3)
hist(Turnovers, main = "Histogram of Turnovers", xlab = "Turnovers")

#Field Goals and Team points
plot(FieldGoals, Points, main = "Field Goals and Team Points", xlab ="Field Goals", ylab="Team Points")
mod5 <-lm(Points ~ FieldGoals)
abline(mod5, col=3)
hist(FieldGoals, main = "Histogram of Field Goals", xlab = "Field Goals")

#Total Fouls and Team points
plot(Fouls, Points, main = "Total Fouls and Team Points", xlab ="Total Fouls", ylab="Team Points")
mod6 <-lm(Points ~ Fouls)
abline(mod6, col=3)
hist(Fouls, main = "Histogram of Fouls", xlab = "Total Fouls")

#Steals and Team points
plot(Steals, Points, main = "Steals and Team Points", xlab ="Steals", ylab="Team Points")
mod6 <-lm(Points ~ Steals)
abline(mod6, col=3)
hist(Steals, main = "Histogram of Steals", xlab = "Steals")

# Reset Plot Grid
par(mfrow=c(1,1))

# Boxplot and Scatterplot
boxplot(bball, main='Box plots for each parameter')
plot(bball,pch=3, main='Scatter Plots of variables')


```

# Model Selection
```{r}
# Initial Full Model
bball.fullmod = lm(TeamPoints~X3PointShots.+FieldGoals.+Assists+TotalRebounds+Turnovers+TotalFouls+Steals, data = bball)
summary(bball.fullmod)


```

# Anova on Initial Full Model
```{r}
library(car)
anova(bball.fullmod)
Anova(bball.fullmod, type="II")

## After ANOVA's result show that we do not need to drop any variables
```

# Diagnostics On Intial Full Model
```{r}
# Residual Plots
library(ALSM)
residualPlots(bball.fullmod, main='Pearson Residual Plots')

# QQ Norm for model residuals
qqnorm(residuals(bball.fullmod), main="Noramal Q-Q Plot for Residuals of model")
qqline(residuals(bball.fullmod), col='green')
```

# Brown Forsythe Test On Initial Full Model
```{r}
library(onewaytests)

# BF-TESTS
bball$residual = bball.fullmod$residuals

bball$AssistsGroup = cut(bball$Assists,5)
bf.test(residual~AssistsGroup, bball)

bball$X3PointShots.Group = cut(bball$X3PointShots.,5)
bf.test(residual~X3PointShots.Group, bball)

bball$FieldGoals.Group = cut(bball$FieldGoals.,5)
bf.test(residual~FieldGoals.Group, bball)

bball$TotalFoulsGroup = cut(bball$TotalFouls,5)
bf.test(residual~TotalFoulsGroup, bball)

bball$TotalReboundsGroup = cut(bball$TotalRebounds,5)
bf.test(residual~TotalReboundsGroup, bball)

bball$TurnoversGroup = cut(bball$Turnovers,5)
bf.test(residual~TurnoversGroup, bball)

bball$StealsGroup = cut(bball$Steals,5)
bf.test(residual~StealsGroup, bball)
```

# Shapiro on Initial Full Model
```{r}
shapiro.test(bball.fullmod$residual)

### Since shapiro test p-value is smaller than 0.05, 
### we reject the null, concluding that the data is NOT NORMAL.
```


# Transformations
```{r}
TeamPointsnew = log(bball$TeamPoints)
bball$TeamPointsnew = log(bball$TeamPoints)
bball.newmod = lm(TeamPointsnew~X3PointShots.+FieldGoals.+Assists+TotalRebounds+Turnovers+TotalFouls+Steals, data = bball)
```

# Diagnostics on new model
```{r}
library(ALSM)
residualPlots(bball.newmod, main='Pearson Residual Plots on New Model')

# QQ Norm for model residuals
qqnorm(residuals(bball.newmod), main="Noramal Q-Q Plot for Residuals of model")
qqline(residuals(bball.newmod), col='green')
```

# Shapiro on new model
```{r}
shapiro.test(residuals(bball.newmod))
```

# Brown-Forsythe Test on new model
```{r}
library(onewaytests)

# BF-TESTS
bball$residual = bball.newmod$residuals

bball$AssistsGroup = cut(bball$Assists,5)
bf.test(residual~AssistsGroup, bball)

bball$X3PointShots.Group = cut(bball$X3PointShots.,5)
bf.test(residual~X3PointShots.Group, bball)

bball$FieldGoals.Group = cut(bball$FieldGoals.,5)
bf.test(residual~FieldGoals.Group, bball)

bball$TotalFoulsGroup = cut(bball$TotalFouls,5)
bf.test(residual~TotalFoulsGroup, bball)

bball$TotalReboundsGroup = cut(bball$TotalRebounds,5)
bf.test(residual~TotalReboundsGroup, bball)

bball$TurnoversGroup = cut(bball$Turnovers,5)
bf.test(residual~TurnoversGroup, bball)

bball$StealsGroup = cut(bball$Steals,5)
bf.test(residual~StealsGroup, bball)
```


# Added Variable Plots
```{r}
avPlots(bball.newmod)
```

# VIF for New Model
```{r}
library(fmsb)
VIF(bball.newmod)
VIF(lm(TeamPointsnew~X3PointShots., data=bball))
VIF(lm(TeamPointsnew~FieldGoals., data=bball))
VIF(lm(TeamPointsnew~Assists, data=bball))
VIF(lm(TeamPointsnew~TotalRebounds, data=bball))
VIF(lm(TeamPointsnew~Turnovers, data=bball))
VIF(lm(TeamPointsnew~TotalFouls, data=bball))
VIF(lm(TeamPointsnew~Steals, data=bball))

VIF(lm(TeamPointsnew~FieldGoals.+X3PointShots.+Assists+Turnovers+TotalFouls+TotalRebounds+Steals, data=bball))

VIF(lm(TeamPointsnew~Assists+FieldGoals.+X3PointShots., data=bball))
VIF(lm(TeamPointsnew~FieldGoals.+X3PointShots.+Assists, data=bball))
VIF(lm(TeamPointsnew~X3PointShots.+Assists+FieldGoals., data=bball))
VIF(lm(TeamPointsnew~X3PointShots.+FieldGoals.+Assists, data=bball))

```


# Outlier Tests
```{r}
# Visuzaliztion for Outliers
# Studentized deleted residuals
influencePlot(bball.newmod, main="Influence Plot")

print('T value for studRes')
qt(1-0.05/(2*328), 328-1-8)

print('Bound for bound hati')
# Hati (FOUND Point )
hati = lm.influence(bball.newmod)$hat
bound = 2*mean(hati)
bound



#Cook threshold
print('Cooks threshold')
qf(0.5,8,320)  #major
```

# Ridge Regression
```{r}
# install.packages("MASS")
# install.packages("lmridge")

# Ridge
library(MASS)
bball.ridgelm = lm.ridge(TeamPointsnew~X3PointShots.+FieldGoals.+Assists+TotalRebounds+Turnovers+TotalFouls+Steals, data = bball, lambda=seq(0,10,0.02))
plot(bball.ridgelm, main="Ridge Trace Plot")
select(bball.ridgelm)

library(lmridge)
bball.ridgemod2 = lmridge(TeamPointsnew~X3PointShots.+FieldGoals.+Assists+TotalRebounds+Turnovers+TotalFouls+Steals, data = bball, K=seq(0,1,0.02))
plot(bball.ridgemod2)
vif(bball.ridgemod2)
```

# Best Sub
```{r}
bs = BestSub(bball[,2:8],bball$TeamPointsnew, num=1)
bs
```

# Cross Validation
```{r}
# Cross validaion
library(leaps)
library(caret)
train.control<-trainControl(method="cv", number=5)
step.model1<-train(TeamPointsnew~X3PointShots.+FieldGoals.+Assists+TotalRebounds+Turnovers+TotalFouls+Steals, data=bball, method="leapBackward",
                   tuneGrid=data.frame(nvmax=7),
                   trControl=train.control)

step.model1$results
```

# ANOVA for New Model
```{r}
bball.newmod
Anova(bball.newmod)
anova(bball.newmod)

# Summary new model
summary(bball.newmod)
```

